March Madness Basketball Turniment

### 2021-04-13 slides/slides.Rmd
added files changed in meeting

---
title: "March Madness Basketball Turniment"
author: "Sabrina Ball and Joe Despres"
date: "updated: `r Sys.Date()`"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

@@

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

@@

https://github.com/EvaMaeRey/flipbookr

@@

>>>>

---
title: "March Madness Basketball Turniment"
author: "Sabrina Ball and Joe Despres"
date: "updated: `r Sys.Date()`"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 6, echo = FALSE, message = FALSE, warning = FALSE, comment = "", cache = F)

library(tidyverse)
library(kableExtra)
library(patchwork)

merged <- readr::read_csv(here::here("data", "merged.csv"))
ids <- readr::read_csv(here::here("data", "ids.csv"))
s2021 <- readr::read_csv(here::here("data", "s2021.csv"))
source(here::here("R", "fitting.R"))
```

@@

@@

# Backround

--

- First tournament was in 1939 only 8 teams

--

- Currently, the top 64 teams qualify

--

- Only one team has ever gone undefeated thought a season and the tournament

--

- 16.9 Million Viewers

--

- MSU won in 1979 and 2000

--

- maybe 2022...

--

- When the turiment starts we form small groups and gamble on the outcomes. Usually by selecting winners and loser from the backet until we have a completely filled out turniment

--

- Nobody has ever made a perfect bracket there are $2^{63}$ possible outcomes.

@@

![](images/empty_bracket.jpg)

---
class: center, middle
# Zoomed in

![](images/zoomed.png)

class: inverse, center, middle

# Can we use basketball seasion data to predict this turniment outcomes?

# Predictors

```{r}
descriptive_statistics <- function(x){
  output <- c(mean(x), median(x), sd(x), min(x), max(x))
  output <- round(output, digits = 2)
  return(output)
}

desc_stats <- merged %>%
  select(fg_percent, ft_percent, rpg,st, to, bkpg) %>%
  map_df(descriptive_statistics) %>%
  t()

colnames(desc_stats) <- c("Mean", "Median", "Std. Dev", "Min", "Max")
rownames(desc_stats) <- c("Field Goal Percentage", "Free Throw Percentage", "Rebounds Per Game", "Steals", "Turnovers", "Blocked Shots Per Game")

kable(desc_stats) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

seasion_seasion <- merged %>%
  nrow()

# n teams
unique_teams <- length(unique(merged$team_id))

# 64 teams in the turniment.

cbind(
    `Games Played` = seasion_seasion,
      Teams = unique_teams,
      Turniment = 64) %>%
  kable() %>%
  kable_styling(bootstrap_options = "striped",
                full_width = F,
                position = "center")
```

---
class: center

# Distribution of Predictors

```{r fig.width=15, fig.height=10}
hists <- merged %>%
  select(x3fg, fg_percent, rpg, st, to,  bkpg)

histogram <- function(x, t){
  ggplot(data = hists, aes(x)) +
    geom_histogram(bins = 50, color = "black", fill = "gray80") +
    theme_light(base_size = 20) +
    labs(title = paste0("", t), x = "", y = "") +
    NULL

}

a <- histogram(merged$x3fg,       t = "Three-Points Goals")
b <- histogram(merged$fg_percent, t = "Percentage of Goals Scored")
c <- histogram(merged$rpg,        t = "Rebounds Per Game")
d <- histogram(merged$st,         t = "Steals Per Game")
e <- histogram(merged$to,         t = "Turnovers Per Game")
f <- histogram(merged$bkpg,       t = "Blocks Per game")
# (a + b + c) / (d + e + f)
(a + b) / (c + d) /(e + f)
```

class: center

# Logistic Model

```{r}
equatiomatic::extract_eq(model = logistic_fit,
                         use_coefs = FALSE,
                         show_distribution = TRUE,
                         wrap = TRUE,
                         terms_per_line = 2,
                         intercept = "beta")
```

# Poisson Model

```{r}
equatiomatic::extract_eq(model = fit_poisson,
                         use_coefs = FALSE,
                         show_distribution = TRUE,
                         wrap = TRUE,
                         terms_per_line = 2,
                         intercept = "beta")
```

# Multinomeal Model

---
class: center 
### Output and Interpretation

```{r}
table_output <- function(obj, rows){
  broom::tidy(obj) %>% 
  mutate(p.value = as.character(p.value)) %>%
  mutate_if(is.numeric, round, 3) %>% 
  mutate(p.value = as.numeric(p.value))  %>% 
  kable(col.names =  c("Term", "Estimate", "Std. Error", "T-Stat", "P-Value")) %>% 
  kable_styling(font_size = 15) %>% 
  row_spec(rows, color = "white", background = "#025cba")
}

table_output(obj = logistic_fit, rows = 2:3)
```

---
class: center 
## Output and Interpretation

```{r}
table_output(obj = logistic_fit, rows = 4:5)
```

---
class: center 
## Output and Interpretation

```{r}
table_output(obj = logistic_fit, rows = 6:7)
```

---
class: center 
## Output and Interpretation

```{r}
table_output(obj = logistic_fit, rows = 8:9)
```

---
class: center 
## Output and Interpretation

```{r}
table_output(obj = logistic_fit, rows = 10:11)
```

---
class: center 
## Output and Interpretation

```{r}
table_output(obj = logistic_fit, rows = 12:13)
```

---
class: center 
## Output and Interpretation

```{r}
table_output(obj = logistic_fit, rows = 14:15)
```

---
class: center
# Scoring

$$\LARGE\sum_{i=1}^{63} (p_i * \mathds{1}_{correct prediction}  -p_i * \mathds{1}_{incorrect perdiction}) = score$$

# Performance

# Conclusion
